Uncovering underlying high-level musical content in the time domain \\ \small Leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn deep audio embeddings with applications to boundary detection tasks

### 2023-06-30 discussion/discussion.tex
Updates from Overleaf

While these do decode underlying high-level musical content effectively remain scientifically unanswered, the outstanding performance in precision metric

@@

>>>>

While whether our model decodes underlying high-level musical content effectively remains scientifically unanswered, the outstanding performance in precision metric

@@

Too little model for raw audio input? \cite{verydeep}

### 2023-06-30 evaluation/evaluation.tex
Updates from Overleaf

Below is a detailed explanation of how we compute the \textit{embeddiogram} from a given audio signal of length $N$. This process comprises loading the audio data, slicing the audio data into windowed segments, processing each window using our pre-trained model to produce a vector per window/time-frame, collecting and stacking these embeddings, and finally normalizing the resulting matrix or as we like to call it, the \textit{embeddiogram}.

@@

\item \textbf{Slice the audio data}: The audio data is segmented into overlapping windows. Each window contains $w$ samples, and consecutive windows are separated by $h$. This gives a total of $H$ windows, defined as:
\begin{equation}
H = 1 + \left\lfloor \frac{N - w}{h} \right\rfloor
\end{equation}
In the case where $\left( N - w \right) \mod h > 0$, we have $H += 1$ to account for the final, potentially smaller, window.

@@

@@

  \begin{minipage}[b]{1.0\linewidth}
    \centering
    \includegraphics[width=\linewidth]{figures/images/355novelty.png}
    \caption[Novelty curve + peaks. Track 391 (SALAMI dataset).]{Novelty curve + peaks. Track 355 (SALAMI dataset).}
    \label{fig:novelty}
  \end{minipage}
\end{figure}

\begin{figure}[ht]
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355recurrencematrix.png} % first figure itself
        \caption[Track 355 (SALAMI dataset) Self-similarity matrix]{Self-similarity matrix computation of the embeddiogram. Track 355 (SALAMI dataset).}
        \label{fig:SSM}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355lagmatrixgaussiansmoothing.png} % second figure itself
        \caption[Track 355 (SALAMI dataset) Self-similarity lag matrix computation of the embeddiogram]{Self-similarity lag matrix. Track 355 (SALAMI dataset).}
        \label{fig:SSLM}
    \end{minipage}
    \vfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355cumulativematrix.png} % third figure itself
        \caption[Track 355 (SALAMI dataset). Q-matrix]{Cumulative matrix computation of the embeddiogram. Track 355 (SALAMI dataset).}
        \label{fig:Q}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355transitivebinarysimilaritymatrix.png} % fourth figure itself
        \caption[Track 355 (SALAMI dataset). Transitive Binary Similarity Matrix]{Transitive Binary Matrix computation of the embeddiogram. Track 355 (SALAMI dataset).}
        \label{fig:TBSM}
    \end{minipage}
\end{figure}

@@

Furthermore, these embeddiograms are a foundation for current state-of-the-art methods, which are expected to receive as input standard audio features. Consequently, they can be processed and manipulated like conventional features \ref{fig:SSM} \ref{fig:SSLM} \ref{fig:Q} \ref{fig:TBSM} \ref{fig:novelty}. Employing such embeddings in a traditional music segmentation algorithm can achieve state-of-the-art performance \cite{deepfeaturesegment}.

@@

While the usefulness of deep audio embeddings can be evaluated in countless downstream tasks \cite{Li2023MERT:Training, Kim2020OneStrategies}, we have chosen music boundary detection, a subset of music structure analysis (MSA) for its popularity \cite{Smith2013ATask}, complexity, and product-compelling nature. This interdisciplinary field aims to understand the structure of music \cite{Nieto2020Audio-BasedApplications}. Due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces challenges like boundary placement ambiguity and similarity quantification \cite{NietoPerceptualMusic}.

All the evaluation experiments have been wrapped around MSAF \cite{MSAF}, a popular framework for MSA.

>>>>

Below is a detailed explanation of how we compute the \textit{embeddiogram} from a given audio signal of length $N$. This process comprises loading the audio data, slicing the audio data into windowed segments, processing each window using our pre-trained model to produce a vector per window/time frame, collecting and stacking these embeddings, and finally, normalizing the resulting matrix.

@@

\item \textbf{Slice the audio data}: The audio data is segmented into overlapping windows. Each window contains $w$ samples, and consecutive windows are separated by a hop size $h$. This gives a total of $H$ windows, defined as:
\begin{equation}
H = 1 + \left\lfloor \frac{N - w}{h} \right\rfloor
\end{equation}
In the case where $\left( N - w \right) \mod h > 0$, we have $H += 1$ to account for the final, potentially smaller window.

@@

We find this to be a good starting compromise solution, allowing enough time span to capture musical content while not being so large that downsampling reduces the information to an unintelligible vector.

@@

  \begin{minipage}[b]{1.0\linewidth}
    \centering
    \includegraphics[width=\linewidth]{figures/images/355novelty.png}
    \caption[Novelty curve + peaks. Track 391 (SALAMI dataset).]{Novelty curve and peak detection. Track 355 (SALAMI dataset).}
    \label{fig:novelty}
  \end{minipage}
\end{figure}

\begin{figure}[ht]
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355recurrencematrix.png} % first figure itself
        \caption[Track 355 (SALAMI dataset) Self-similarity matrix]{Self-similarity matrix computation of the embeddiogram. Track 355 (SALAMI dataset).}
        \label{fig:SSM}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355lagmatrixgaussiansmoothing.png} % second figure itself
        \caption[Track 355 (SALAMI dataset) SSLM computation of the embeddiogram]{Self-similarity lag matrix. Track 355 (SALAMI dataset).}
        \label{fig:SSLM}
    \end{minipage}
    \vfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355cumulativematrix.png} % third figure itself
        \caption[Track 355 (SALAMI dataset). Q-matrix]{Cumulative matrix computation of the embeddiogram. Track 355 (SALAMI dataset) .}
        \label{fig:Q}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355transitivebinarysimilaritymatrix.png} % fourth figure itself
        \caption[Track 355 (SALAMI dataset). Transitive Binary Similarity Matrix]{Transitive Binary Matrix computation of the embeddiogram. Track 355 (SALAMI dataset).}
        \label{fig:TBSM}
    \end{minipage}
\end{figure}

@@

These deep audio features can be a foundation for current state-of-the-art methods, which are expected to receive as input standard traditional audio features. Consequently, they can be processed and manipulated like conventional features as described by \cite{unsuperMSA} and displayed in \ref{fig:SSM}, \ref{fig:SSLM}, \ref{fig:Q}, and \ref{fig:TBSM} \ref{fig:novelty}. Employing such embeddings in a traditional music segmentation algorithm can achieve state-of-the-art performance \cite{deepfeaturesegment}.

@@

While the usefulness of deep audio embeddings can be evaluated in countless downstream tasks \cite{Li2023MERT:Training, Kim2020OneStrategies}, we have chosen music boundary detection (also known as track segmentation), a subset of music structure analysis (MSA) for its popularity \cite{Smith2013ATask}, complexity, and product-compelling nature. This interdisciplinary field aims to understand the structure of music \cite{Nieto2020Audio-BasedApplications}. Due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces non-solved challenges like boundary placement ambiguity and similarity quantification \cite{NietoPerceptualMusic}.

### 2023-06-30 future work/future work.tex
Updates from Overleaf

The question as per \cite{Turian2022HEAR:Representations} or \cite{Li2023MERT:Training} remains if such general-purpose audio representation can mimic human hearing. That is a "thorn in our side" that we will pursue.

@@

Utilize the dB-scale Mel-spectrum magnitude of audio as input data; it is a popular choice for input representation in music-related tasks when applying CNNs, as has been demonstrated in various studies \cite{Kim2020OneStrategies}. While raw audio is the most meaningful audio representation, it's been reported that dB-scale Mel-spectrum frequency-domain summarization, grounded in psycho-acoustics, is computationally efficient and challenging to replicate solely through data-driven methods \cite{Kim2020OneStrategies}; therefore, a trade-off that looks worthy of being explored.

>>>>

@@

Utilize the dB-scale Mel-spectrum magnitude and CQT of audio as input data; it is a popular choice for input representation in music-related tasks when applying CNNs, as has been demonstrated in various studies \cite{Kim2020OneStrategies}. While raw audio is the most meaningful audio representation, it's been reported that dB-scale Mel-spectrum frequency-domain summarization, grounded in psycho-acoustics, is computationally efficient and challenging to replicate solely through data-driven methods \cite{Kim2020OneStrategies}; therefore, a trade-off that looks worthy of being explored.

### 2023-06-30 introduction/Subsections/Introduction.tex
Updates from Overleaf

\begin{itemize}
\item Creating an audio embedding method that excels across numerous applications without the necessity for fine-tuning \cite{Turian2022HEAR:Representations}
\item Boosting the classification of environmental sounds \cite{Kim2020OneStrategies, CramerLOOKEMBEDDINGS}
\item Enhancing vocal-centric music tasks through cross-domain audio embeddings \cite{Kim2021LearningLoss}
\item Integrating task-specific and pre-trained features to optimize audio classification \cite{Hung2022Feature-informedClassification}
\item Developing a music similarity search engine for video production \cite{epidemic}
\item Improving Music Emotion Recognition (MER) performance \cite{KohComparisonRecognition}
\item Evaluating the effectiveness of speaker recognition via pre-trained model embeddings \cite{lightweight}
\item Embedding songs for artist identification to facilitate similarity comparisons \cite{contentmusicsimtriplet2020}
\item Addressing cross-modal text-to-music retrieval issues \cite{WonEmotionStories}
\item Enabling automated music rearrangement \cite{Stoller2018IntuitiveTransitions, Plachouras2023MusicSegmentation}
\end{itemize}

@@

Given the proven effectiveness of deep audio embeddings in existing research and their potential for transfer learning, they hold considerable promise for MIR downstream tasks, including boundary detection. However, it is yet to be determined whether a general-purpose audio representation can successfully emulate human hearing \cite{Turian2022HEAR:Representations}, even though some techniques have demonstrated generalizability across many music-understanding tasks \cite{Li2023MERT:Training}.

Yet, despite these strides, deep learning's role in music understanding remains relatively nascent, with scarce work on deep music representations, a dearth of large-scale datasets, and a lack of a universal community-driven benchmark \cite{Yuan2023MARBLE:Evaluation}.

>>>>

\begin{itemize}
\item Creating an audio embedding method that excels across numerous applications without the necessity for fine-tuning \cite{Turian2022HEAR:Representations}
\item Boosting the classification of environmental sounds \cite{CramerLOOKEMBEDDINGS}
\item Enhancing vocal-centric music tasks through cross-domain audio embeddings \cite{Kim2021LearningLoss}
\item Integrating task-specific and pre-trained features to optimize audio classification \cite{Hung2022Feature-informedClassification}
\item Developing a music similarity search engine for video production \cite{epidemic}
\item Improving Music Emotion Recognition (MER) performance \cite{KohComparisonRecognition}
\item Evaluating the effectiveness of speaker recognition via pre-trained model embeddings \cite{lightweight}
\item Embedding songs for artist identification to facilitate similarity comparisons \cite{contentmusicsimtriplet2020}
\item Addressing cross-modal text-to-music retrieval issues \cite{WonEmotionStories}
\item Enabling automated music rearrangement \cite{Stoller2018IntuitiveTransitions, Plachouras2023MusicSegmentation}
\end{itemize}

@@

Given the proven effectiveness of deep audio embeddings in existing research and their potential for transfer learning, they hold considerable promise for MIR downstream tasks, including boundary detection. However, it is yet to be determined whether a general-purpose audio representation can successfully emulate human hearing \cite{Turian2022HEAR:Representations}, even though some techniques have demonstrated generalizability across many music-understanding tasks \cite{Li2023MERT:Training, Kim2020OneStrategies}.

Yet, despite these strides, deep learning's role in music understanding remains relatively nascent, with scarce work on deep music representations, a dearth of large-scale datasets, and a lack of a universal community-driven benchmark \cite{Turian2022HEAR:Representations, Yuan2023MARBLE:Evaluation}.

### 2023-06-30 methods/subsections/deep learning.tex
Updates from Overleaf

Deep learning (DL), a subset of machine learning, employs deep neural networks (DNNs) to decipher hierarchical data representations, allowing models with multiple processing layers to abstract data across numerous levels \cite{LeCun2015DeepLearning}. DL models have significantly improved various domains due to their nonlinear modeling capabilities and scalability with large datasets.

>>>>

Deep learning (DL), a subset of machine learning, employs deep neural networks (DNNs) to decipher hierarchical data representations, allowing models with multiple processing layers to abstract data across numerous levels \cite{LeCun2015DeepLearning, Sarker2021DeepDirections}. DL models have significantly improved various domains due to their nonlinear modeling capabilities and scalability with large datasets.

### 2023-06-30 methods/subsections/implementation.tex
Updates from Overleaf

Our implementation is a slight tweak to the \textit{contrastive learning of musical representations} (CLMR), a method that learns valuable, discriminative music representations without explicit labels by contrasting positive augmentations of a musical piece against negative ones \cite{CLMR2021}. CLMR falls under a subset of ML called self-supervised learning (SSL) \cite{Balestriero2023ALearning}, where models learn independently from unlabeled data to create their supervisory signals \cite{audioselfsupsurvey}. This method resembles how humans learn from observations and interactions, turning unsupervised problems into supervised ones by auto-generating labels. Benefits of SSL include reduced dependence on labeled data, which contributes to more robust and generalizable data representations.

As stated above, the methodology of CLMR involves generating different augmentations of the same musical piece to serve as positive pairs while using other pieces as antagonistic pairs. The training process encourages the model to produce similar representations for positive and dissimilar ones for antagonistic pairs.

@@

This work suggests employing a Triple Siamese Network, a model architecture known for its efficacy in music similarity retrieval tasks \cite{contentmusicsimtriplet2020}. The aim is to minimize the loss function between a triplet of anchor, positive, and negative samples, utilizing online triplet mining for optimizing memory resources \cite{Sikaroudi2020OfflinePatches}. This SSL model aims to distinguish between similar and dissimilar samples effectively.

Introduced by Bromley and LeCun \cite{Bromley1993SignatureNetwork}, Siamese Networks are deep learning architectures designed for tasks requiring comparison or similarity assessment between instances. The architecture comprises identical subnetworks that share the same parameters, improving memory usage and computational efficiency. Rather than learning specific features of individual classes, they focus on a similarity metric, making them ideal for imbalanced datasets. Each subnetwork processes an input independently, combining the outputs to yield a similarity score. Training with shared weights enables the model to learn invariant input representation, improving comparison efficiency. This is accomplished through specialized loss functions such as contrastive or triplet loss, aiming to minimize the distance between similar inputs and maximize it for dissimilar ones.

The Triple Siamese Network extends a binary contrastive architecture by comparing three input instances instead of two. It aims to learn an embedding space wherein similar instances are closer and dissimilar ones are distant. The inputs, consisting of an anchor, a positive instance (similar to the anchor), and a negative instance (dissimilar to the anchor), are processed by identical subnetworks sharing the same parameters.

@@

With 2.4 million trainable parameters, this fully convolutional model reduces computational requirements and learns features at different scales through its multi-resolution architecture \ref{tab:samplecnntable}.

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{>{\hsize=1.6\hsize}X>{\hsize=0.6\hsize}X>{\hsize=0.6\hsize}X}
\toprule
\thead{\textbf{Layer Type}} & \thead{\textbf{In Channels}} & \thead{\textbf{Out Channels}} \\
\midrule
Conv + ReLU & 1 & 128 \\
\addlinespace
\multicolumn{3}{c}{The following layers are repeated depending on the strides and hidden parameters:} \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 128 & 128 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 128 & 128 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 128 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 512 \\
\addlinespace
Conv + BatchNorm + ReLU + AvgPool & 512 & 512 \\
\addlinespace
Fully connected or representation layer & 512 & 128 \\
\bottomrule
\end{tabularx}
\caption{Layer specifications for SampleCNN model}
\label{tab:samplecnntable}
\end{table}

The original model has been modified slightly to serve our specific needs: the output from the final convolutional layer is now subjected to an average pooling operation.

@@

Previous research has utilized CNNs with various features such as Mel-Scaled Log-magnitude Spectograms (MLS), Self-Similarity Matrices (SSM), and Self-Similarity Lag Matrices (SSLM) as inputs \cite{Hernandez-Olivan2021MusicFeatures}. However, features derived from raw audio may lack interpretability in some scenarios \cite{Schindler2020DeepTutorial}, and raw audio presents unique advantages despite being computationally demanding. It ensures the preservation of the original signal, potentially uncovering novel insights, and allows for direct feature extraction via advanced DL models \cite{learning}. Nevertheless, it comes with challenges, such as high dimensionality requiring substantial computational resources. Moreover, it is sensitive to noise, adding complexity to the extraction process.

Time-domain processing naturally handles temporal patterns and sequences in data, thereby avoiding windowing artifacts. Although audio feature-based methods are effective for various audio-related machine learning tasks, their limitations lie in representing perceptual similarity. Mel-Scaled Log-magnitude Spectograms, for example, capture frequency distribution over time. Yet, the complexity of human auditory perception—encompassing temporal patterns, phase relationships between frequencies, and higher-level musical structures—means that musically similar sounds can have distinct spectrograms. This discrepancy implies that using spectrogram distance alone for measuring high-level music content may not always align with human perceptions \cite{Kim2020OneStrategies, Mesostructures2023}.

@@

While we have experimented with helpful audio augmentation tools \cite{Spijkervet2021Spijkervet/torchaudio-augmentations:V1.0, Kharitonov2020DataDomain}, the specific requirements of our experiments required the development of our transformation chain using \textit{torchaudio}'s \cite{Yang2021TorchAudio:Processing} implementation of \textit{SoX} \cite{sox}: given an anchor audio signal $A[n]$, we generate a positive signal $P[n]$ by applying a series of amplitude, time-domain, frequency-domain, modulation, reverberation, and nonlinear effects with additive noise on top of it.

@@

\textbf{Time-domain effects}: The signal's playback speed and duration are altered through speed change and stretching, preserving the perceptual musical relationships. The respective factors are $\alpha \in [0.9, 1.1]$ for speed change and $\beta \in [0.9, 1.1]$ for stretch.

@@

\textbf{Modulation effects}: Effects such as chorus and tremolo employ a control signal or low-frequency oscillator. The specified ranges determine the chorus parameters, while the tremolo's amplitude modulation frequency and depth are controlled by $f_m \in [0.1, 100]$ and $d_m \in [1, 101]$, respectively.

\textbf{Reverberation effects}: These effects simulate a physical space's acoustic reflections and reverberations by applying an impulse response $h_R[n]$.

@@

\begin{equation}\label{eq:positive_signal}
P[n] = A[n] \ast h_{G}[n] \ast h_{C}[n] \ast h_{D}[n] \ast h_{P_t}[n] \ast h_{R}[n] \ast h_{S}[n] \ast h_{T}[n] \ast h_{T_m}[n] + Noise[n]
\end{equation}

The $\ast$ symbol represents convolution, and the various $h$ symbols indicate the impulse responses of specific effects. The noise signal $N[n]$ is included with a specific SNR.

@@

\begin{enumerate}
\item We first calculate the minimum and maximum audio chunk lengths in samples:
\begin{equation}
l_{min} = t_{min} \times S
\end{equation}
\begin{equation}
l_{max} = t_{max} \times S
\end{equation}

@@

The whole purpose of this process is to disturb the content unfolding in the time domain.

@@

On the contrary of the also popular contrastive loss \cite{supercontrast}, the triplet loss function directs the learning process by minimizing the distance between the anchor and positive instances \textbf{and} maximizing the distance between the anchor and negative instances. Including a margin parameter in the loss function guarantees a minimum separation between positive and negative instances in the embedding space.

@@

The deep learning models were trained on a high-performance cloud computing setup hosted on \href{https://cloud.google.com/}{Google Cloud Platform}. The machine was of type \texttt{n1-standard-32}, equipped with an Intel Skylake processor and four NVIDIA Tesla T4 GPUs. The multiple GPUs allowed for efficient parallel processing, significantly reducing the training time for the deep learning models.

>>>>

Our implementation offers a slight modification to the \textit{contrastive learning of musical representations} (CLMR), a method that learns valuable, discriminative music representations without explicit labels by contrasting positive augmentations of a musical piece against negative ones \cite{CLMR2021}. CLMR falls under a branch of ML known as self-supervised learning (SSL) \cite{Balestriero2023ALearning}, where models learn autonomously from unlabeled data to create their own supervisory signals \cite{audioselfsupsurvey}. This method resembles how humans learn from observations and interactions, transforming unsupervised problems into supervised ones by auto-generating labels. SSL benefits include reduced dependence on labeled data, contributing to more robust and generalizable data representations.

As previously stated, the methodology of CLMR involves generating different augmentations of the same musical piece to serve as positive pairs while using other pieces as negative pairs. The training process encourages the model to produce similar representations for positive and dissimilar ones for negative pairs.

@@

This work suggests employing a Triple Siamese Network (TSN), a model architecture known for its efficacy in music similarity retrieval tasks \cite{contentmusicsimtriplet2020}. The aim is to minimize the loss function between a triplet of anchor, positive, and negative samples, utilizing online triplet mining for optimizing memory resources \cite{Sikaroudi2020OfflinePatches}. This SSL model aims to distinguish between similar and dissimilar samples effectively.

Introduced by Bromley and LeCun \cite{Bromley1993SignatureNetwork}, Siamese Networks are deep learning architectures designed for tasks requiring comparison or similarity assessment between instances. The architecture comprises identical subnetworks that share the same parameters, improving memory usage and computational efficiency. Rather than learning specific features of individual classes, they focus on a similarity metric, making them ideal for imbalanced datasets. Each subnetwork processes an input independently, combining the outputs to yield a similarity score. Training with shared weights enables the model to learn invariant input representation, improving comparison efficiency. This is accomplished through a specialized loss function called triplet loss, which aims to minimize the distance between similar inputs and maximize it for dissimilar ones.

The TSN extends the binary contrastive architecture by comparing three input instances instead of two. It strives to learn an embedding space where similar instances are closer and dissimilar ones are more distant. The inputs, consisting of an anchor, a positive instance (similar to the anchor), and a negative instance (dissimilar to the anchor), are processed by identical subnetworks sharing the same parameters.

@@

With only 2.4 million trainable parameters, this fully convolutional model reduces computational requirements and learns features at different scales through its multi-resolution architecture \ref{tab:samplecnntable}.

\begin{table}[ht]
\centering
\small
\begin{tabularx}{\textwidth}{>{\hsize=1.6\hsize}X>{\hsize=0.6\hsize}X>{\hsize=0.6\hsize}X}
\toprule
\thead{\textbf{Layer Type}} & \thead{\textbf{In Channels}} & \thead{\textbf{Out Channels}} \\
\midrule
Conv + ReLU & 1 & 128 \\
\addlinespace
\multicolumn{3}{c}{The following layers are repeated depending on the strides and hidden parameters:} \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 128 & 128 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 128 & 128 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 128 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 256 \\
\addlinespace
Conv + BatchNorm + ReLU + MaxPool & 256 & 512 \\
\addlinespace
Conv + BatchNorm + ReLU + AvgPool & 512 & 512 \\
\addlinespace
Fully connected or representation layer & 512 & 128 \\
\bottomrule
\end{tabularx}
\caption{Layer specifications for SampleCNN model}
\label{tab:samplecnntable}
\end{table}

The original model has been subtly adjusted by introducing an average pooling operation to the final convolutional layer. This modification, dimensionality reduction, is strategic for handling various input data sizes. Condensing the feature maps to a fixed-size output ensures consistency for subsequent layers like fully connected ones, streamlining computations while maintaining crucial information. This results in a more robust, efficient learning process that offers translation invariance and better contextual understanding.

@@

Previous research has utilized CNNs with various features such as Mel-Scaled Log-magnitude Spectograms (MLS), Self-Similarity Matrices (SSM), and Self-Similarity Lag Matrices (SSLM) as inputs \cite{Hernandez-Olivan2021MusicFeatures}. However, features derived from raw audio may lack interpretability in some scenarios \cite{Schindler2020DeepTutorial}, and raw audio presents unique advantages despite being highly computationally demanding. It ensures the preservation of the original signal, potentially uncovering novel insights, and allows for direct feature extraction via advanced DL models \cite{learning, verydeep}. Nevertheless, it comes with challenges, such as high dimensionality requiring substantial computational resources.

Time-domain processing naturally handles temporal patterns and sequences in data, thereby avoiding windowing artifacts. Although audio feature-based methods are effective for various audio-related machine learning tasks, their limitations lie in representing perceptual similarity. Mel-Scaled Log-magnitude Spectograms, for example, capture frequency distribution over time. Yet, the complexity of human auditory perception—encompassing temporal patterns, phase relationships between frequencies, and higher-level musical structures means that musically similar sounds can have distinct spectrograms. This discrepancy implies that using spectrogram distance alone for measuring high-level music content may not always align with human perceptions \cite{Kim2020OneStrategies, Mesostructures2023}.

@@

While we have experimented with helpful audio augmentation tools such \cite{Spijkervet2021Spijkervet/torchaudio-augmentations:V1.0, Kharitonov2020DataDomain}, the specific requirements of our experiments required the development of our own transformation chain using \textit{torchaudio}'s \cite{Yang2021TorchAudio:Processing} implementation of \textit{SoX} \cite{sox}: given an anchor audio signal $A[n]$, we generate a positive signal $P[n]$ by applying a series of amplitude, time-domain, frequency-domain, modulation, reverberation, and nonlinear effects with additive noise on top of it.

@@

\textbf{Time-domain effects}: The signal's playback speed and duration are altered through speed change and stretching, preserving the relative perceptual musical relationships between wave points. The respective factors are $\alpha \in [0.9, 1.1]$ for speed change and $\beta \in [0.9, 1.1]$ for stretch.

@@

\textbf{Modulation effects}: Utilize a control signal or low-frequency oscillator. The chorus parameters are determined by a set of six different variables: $f_d, f_m, f_b, f_w, f_a, f_s$ where $f_d, f_m, f_w, f_a$ are floats in the range $[0.1, 1.0]$, $f_b$ is an integer in the range $[20, 55]$, $f_s$ is an integer in the range $[2, 5]$ and the type of control signal is chosen between "s" and "t". The tremolo's amplitude modulation frequency and depth are controlled by $t_s \in [0.1, 100]$ and $t_d \in [1, 101]$, respectively.

\textbf{Reverberation effects} simulate a physical space's acoustic reflections and reverberations by applying an impulse response $h_R[n]$.

@@

\begin{equation}\label{eq:positive_signal}
P[n] = A[n] \ast h_{G}[n] \ast h_{C}[n] \ast h_{D}[n] \ast h_{P_t}[n] \ast h_{S}[n] \ast h_{T}[n] + N_{SNR}[n]
\end{equation}

The $\ast$ symbol represents convolution, and the various $h$ symbols indicate the impulse responses of specific effects. The noise signals $N_{SNR}[n]$ are included with a specific SNR.

@@

\begin{enumerate}
\item We first calculate the minimum and maximum audio chunk lengths in samples:
\begin{align}
l_{min} &= t_{min} \times S, &
l_{max} &= t_{max} \times S
\end{align}

@@

The whole purpose of this process is to disturb the content unfolding in the time domain so it becomes musically unintelligible while maintaining the production and sonic attributes.

@@

On the contrary to the also popular contrastive loss \cite{supercontrast}, the triplet loss function directs the learning process by minimizing the distance between the anchor and positive instances \textbf{and} maximizing the distance between the anchor and negative instances. Including a margin parameter in the loss function guarantees a minimum separation between positive and negative instances in the embedding space.

@@

The deep learning models were trained on a high-performance cloud computing setup hosted on \href{https://cloud.google.com/}{Google Cloud Platform}. The machine was of type \texttt{n1-standard-32}, equipped with an Intel Skylake processor and four NVIDIA Tesla T4 GPUs. The multiple GPUs allowed for efficient parallel processing, significantly reducing the training time.

### 2023-06-30 results/subsections/tables.tex
Updates from Overleaf

The evaluation results have been obtained using the standard and most commonly used MIR's tools and frameworks, such as default values of MSAF \cite{MSAF} for the segmentation algorithms \cite{sf}, the SALAMI dataset \cite{Smith2011DESIGNANNOTATIONS} as evaluation ground truth and $mir_eval$ \cite{RaffelMir_eval:METRICS} for metric computation.

@@

\begin{table}
  \centering
  \begin{adjustbox}{width=\textwidth}
  \begin{threeparttable}
    \begin{tabular}{c|l|c|c|c} 
\toprule
     \multicolumn{2}{c|}{\textbf{Training dataset} and \textbf{Feature}}  & \multicolumn{1}{c}{\textbf{SF}} & \multicolumn{1}{c}{\textbf{Foote}} & \multicolumn{1}{c}{\textbf{CNMF}} 
     \\ \midrule 
    \multirow{5}{*}{}&   
         \multirow{1}{*}{MFCC } & \usebox{\mfccSF} & \usebox{\mfccFoote} & \usebox{\mfccCNMF} \\\cline{2-5}
        & \multirow{1}{*}{PCP} & \usebox{\pcpSF} & \usebox{\pcpFoote} & \usebox{\pcpCNMF} \\\cline{2-5}
        & \multirow{1}{*}{Tonnetz} & \usebox{\tonnetzSF} & \usebox{\tonnetzFoote} & \usebox{\tonnetzCNMF} \\\cline{2-5} 
        & \multirow{1}{*}{CQT} & \usebox{\cqtSF} & \usebox{\cqtFoote} & \usebox{\cqtCNMF} \\\cline{1-5} 
    \multirow{1}{*}{GTZAN}& 
        \multirow{1}{*}{Embeddiogram} & \usebox{\embeddioaSF} & \usebox{\embeddioaFoote} & \usebox{\embeddioaCNMF} \\\cline{1-5} 
    \multirow{1}{*}{MSD}& 
        \multirow{1}{*}{Embeddiogram} & \usebox{\embeddiobSF} & \usebox{\embeddiobFoote} & \usebox{\embeddiobCNMF} \\\bottomrule
    \end{tabular}
    \caption[Metric comparison for different audio features and different segmentation algorithms]{Comparison of precision, recall, and f-measure for different features obtained using the SF\cite{sf}, Foote \cite{Foote2000AutomaticNovelty}, and CNMF \cite{NietoCONVEXIDENTIFICATION} algorithms on the SALAMI dataset. Sliding windowed segments across the audio signal input signal is 4 seconds}\label{ta:results}
  \end{threeparttable}
  \end{adjustbox}
\end{table}

>>>>

The evaluation results were obtained using standard and commonly used MIR tools and frameworks. These include the MSAF \cite{MSAF} implementation for segmentation algorithms, the SALAMI dataset \cite{Smith2011DESIGNANNOTATIONS} for evaluation ground truth, and $mir-eval$ \cite{RaffelMir_eval:METRICS} for metric computation.

The results are displayed in \ref{ta:results}

@@

\begin{table}
  \centering
  \begin{adjustbox}{width=\textwidth}
  \begin{threeparttable}
    \begin{tabular}{c|l|c|c|c} 
\toprule
     \multicolumn{2}{c|}{\textbf{Training dataset} and \textbf{Feature}}  & \multicolumn{1}{c}{\textbf{SF}} & \multicolumn{1}{c}{\textbf{Foote}} & \multicolumn{1}{c}{\textbf{CNMF}} 
     \\ \midrule 
    \multirow{5}{*}{}&   
         \multirow{1}{*}{MFCC } & \usebox{\mfccSF} & \usebox{\mfccFoote} & \usebox{\mfccCNMF} \\\cline{2-5}
        & \multirow{1}{*}{PCP} & \usebox{\pcpSF} & \usebox{\pcpFoote} & \usebox{\pcpCNMF} \\\cline{2-5}
        & \multirow{1}{*}{Tonnetz} & \usebox{\tonnetzSF} & \usebox{\tonnetzFoote} & \usebox{\tonnetzCNMF} \\\cline{2-5} 
        & \multirow{1}{*}{CQT} & \usebox{\cqtSF} & \usebox{\cqtFoote} & \usebox{\cqtCNMF} \\\cline{1-5} 
    \multirow{1}{*}{GTZAN \cite{GTZAN}}& 
        \multirow{1}{*}{Embeddiogram} & \usebox{\embeddioaSF} & \usebox{\embeddioaFoote} & \usebox{\embeddioaCNMF} \\\cline{1-5} 
    \multirow{1}{*}{MSD \cite{MSD}}& 
        \multirow{1}{*}{Embeddiogram} & \usebox{\embeddiobSF} & \usebox{\embeddiobFoote} & \usebox{\embeddiobCNMF} \\\bottomrule
    \end{tabular}
    \caption[Metric comparison for different audio features and different segmentation algorithms]{Comparison of precision, recall, and f-measure for different features obtained using the SF\cite{sf}, Foote \cite{Foote2000AutomaticNovelty}, and CNMF \cite{NietoCONVEXIDENTIFICATION} algorithms on the SALAMI dataset. Sliding windowed segments across the audio signal input signal is 4 seconds}\label{ta:results}
  \end{threeparttable}
  \end{adjustbox}
\end{table}
